---
title: "Corruption in the lab"
author: "Fedor Anna, Mokos Judit"
date: "19/12/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Description of the experiment

Sequential dyadic die-rolling task
Two-by-two design:
- Game: simple game or charity game
- Partner: simulated honest partner or simulated dishonest partner

Participants are randomly assigned to one of the four conditions:
- SH: simple game with honest partner
- SD: simple game with dishonest partner
- CH: charity game with honest partner
- CD: charity game with dishonest partner

Each participant plays 20 rounds of the game.A round of game consists of the following steps:
- Participant learns the reported number of the supposed partner
- Participant throws a dice
- Participant reports the number 
- Both players get paid according to the reported numbers
- In the charity game a charity foundation gets a small amount of donation

Data from participants: 20 reported numbers, Questionnaire, Finger lenghts, Saliva ???
Predictors: condition, (age, gender), ???

Data frame column names:
Each throw is an observation:
ID
Game
Partner
ValueA (partenr hanyast dobott)
ValueB (hanyast dobott)
Questionnaire ?
Finger ?
Saliva ?

# Figures

## Scatter plots of individual behavior
These should be based on Weisel & Shalvi, 2015, Fig. S10.
One plot per participant. Plots from the same condition should be compiled into one composite figure.
```{r}

```

## Heat map to demonstrate the distribution of reported numbers 
Like Weisel & Shalvi, 2015, Fig. 2, but instead of circles, each rectangle should be color coded according the number of observations within them.
One figure per condition.
```{r}

```

## Box plots of the mean number of doubles
One box for each condition on the same figure
```{r}

```

## Box plots of the mean reports: NINCS ÉRTELME
One box for each condition on the same figure
```{r}

```

# Statistics

## Distribution of reported numbers 
Kolmogorov-Smirnov
One per condition

## The number of doubles
Each participant ("dyad"") is a single observation: the number of reported doubles
Wilcoxon signed-rank U test
Expected value: 3.33 doubles/20 trials (16.7%; 20*1/6)
Separately for each condition
```{r}

```

## Mean reports: VAN ENNEK ÉRTELME NÁLUNK? NINCS
Each participant ("dyad"") is a single observation: the mean of the reported numbers
Wilcoxon signed-rank U test
Expected value: 3.5
Separately for each condition
```{r}

```
## Compare the number of doubles in pairs of conditions
Mann-Whitney U test
```{r}

```

## Effect size
To estimate the effect size of the honest-dishonest treatment let's compare the choices of participant Bs, who were paired with an honest A vs. a chaeting A
Data: from the supplementary material of Weisel&Shalvy, 2015 and from Wouda et al., 2015
```{r}

```

## Linear regression to test the effect of predictors
Dependent variable: number of reported doubles (interval)
Predictors: game (binary), partner (binary), finger length (interval), hormone levels (interval)

## random or non-random, khi square


## LM? 