---
title: "Corruption in the lab"
author: "Fedor Anna, Mokos Judit"
editor_options: null
output:
  word_document: default
  pdf_document: default
  html_document: default
chunk_output_type: console
---
  
```{r setup, include=FALSE}

rm(list=ls())
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(rmarkdown)
library(dplyr)
library(ggplot2)
library(knitr)

#datafile = "C:/Users/fedor/OneDrive/Documents/DOKUMENTUMOK/Corruption.Project/GitHub/corruption_in_the_lab/power_simulations/RESULTS/20191002-102445_dummydata.RData" 
#datafile = "/Users/mokosjudit/Documents/GitHub/corruption_in_the_lab/power_simulations/RESULTS/20191002-102445_dummydata.RData" 

```

<!-- # Description of the experiment 

The game: Sequential dyadic die-rolling task with a simulated partner. Human participants always play the role of "Player B" (the player that rolls the dice second).

Two-by-two design:
  
  * Game: simple game or charity game
  * Partner: simulated honest partner or simulated dishonest partner  

Participants are randomly assigned to one of the four conditions:
  
  * SH: simple game with honest partner
  * SD: simple game with dishonest partner
  * CH: charity game with honest partner
  * CD: charity game with dishonest partner  

Each participant plays 20 rounds of the game. A round of game consists of the following steps:
  
  1. The reported number of the partner appears on the screen
  2. Participant throws a dice privately
  3. Participant reports the number 
  4. Both players get a score according to the reported numbers: the score is the reported value*300, if they reported a double; otherwise it's 0
  5. In the charity game a charity foundation gets a small amount of donation  

# Dummy data: simulated data
Data frame column names (each row is a dice roll):

* ID: random ID of participant (10000:99999)
* Game: simple or charity (S/C)
* Partner: honest or dishonest (H/D)
* Condition: one of the four experimental conditions (SH, SD, CH, CD)
* Index: index of round of game (1:20)
* ValueA: value of simulated dice roll (1:6)
* ValueB: value of participant's reported dice roll (1:6)
* Double: whether the participant reported a double (1/0)
* Q1: answer to first questionnaire quiestion (a random letter)
* Q2: answer to second questionnaire question (a random letter)
* Fingerratio: ratio of two fingers (normal distribution, mean=1, sd=0.1)
-->

```{r Load data, warning=F, error=F, message=F}

#load(datafile)
rawdata <- read.csv("/Users/mokosjudit/Google Drive/Korrupció_a_laborban_ötletek/eles_nevaltoztassrajta/csv/template_datasets/data_ztree_pilot.csv")
 

  a <- count(rawdata, ID, Condition)/20
  tested_samplesizes <- sum(a$n)



parameters <- list(random_seed=NA, 
                   tested_samplesizes=tested_samplesizes,
                   nbof_simulations_per_samplesize=NA,
                   nbof_trials=20, 
                   dice_sides=6,
                   dishonestA_distribution=NA,
                   probof_doubles=NA )


```
# Results

## Participants
```{r Data checking}

participants <- rawdata %>%
  group_by(ID) %>%
  summarize(
    Condition = unique(Condition), 
    Game = unique(Game),
    Partner = unique(Partner),
    Nbof_rounds = max(Index),
    Invalid_values = sum(ValueA<1)+sum(ValueA>7)+sum(ValueB<1)+sum(ValueB>7),
    Nbof_doubles = sum(Double), 
    Avg_report = mean(ValueB),
    Fingerratio = unique(Fingerratio))

excluded_IDs <- c(
  participants$ID[which(participants$Nbof_rounds<20)],
  participants$ID[which(participants$Nbof_rounds>20)], 
  participants$ID[which(participants$Invalid_values>0)]
)
nbof_excluded <- length(excluded_IDs)

rawdata <- filter(rawdata, is.element(ID,excluded_IDs)==FALSE)

participants <- rawdata %>%
  group_by(ID) %>%
  summarize(
    Condition = unique(Condition), 
    Game = unique(Game),
    Partner = unique(Partner),
    Nbof_rounds = max(Index),
    Invalid_values = sum(ValueA<1)+sum(ValueA>7)+sum(ValueB<1)+sum(ValueB>7),
    Nbof_doubles = sum(Double), 
    Avg_report = mean(ValueB),
    Fingerratio = unique(Fingerratio))

groups <- participants %>%
  group_by(Condition) %>%
  arrange(desc(Condition)) %>%
  summarize(
    Nbof_participants = n_distinct(ID),
    Nbof_doubles_per_group = sum(Nbof_doubles),
    Avg_nbof_doubles = Nbof_doubles_per_group/Nbof_participants,
    Median_nbof_doubles = median(Nbof_doubles),
    Percentage_of_doubles = round((Nbof_doubles_per_group/(Nbof_participants*parameters$nbof_trials))*100),
    Avg_avg_report = mean(Avg_report))
groups <- arrange(groups, desc(Condition))

```

We have excluded `r nbof_excluded` participants because XXX. We kept recruiting participants until we had 36 in each condition, all in all 144 participants. Their median age was `r median(rawdata$age)` (range `r min(rawdata$age)`  - `r max(rawdata$age)`) and `r length(which(rawdata$sex==1))/20` were male.

 
## The distribution of reported numbers 
The following shows the frequency of value pairs in the four conditions: Higher frequency pairs are darker. 
*It can be seen that the diagonal (representing doubles) is darker than the rest of the figure in all conditions and that 6-6 pairs are especially dark in conditions with a dishonest partner, as expected.*

```{r Heat maps}

cond <- as.factor(unique(rawdata$Condition))
maximum <- max(count(rawdata, ValueA, ValueB, Condition)$n)

for(i in cond){
  heatmap_df <- data.frame(count(filter(rawdata, Condition==i), ValueA, ValueB))
  
  #yourfilename=paste("figure2_", i, ".jpg",sep="") #filename for sacing
  #jpeg(file=yourfilename)
  
  par(pty="s")
  figure2 <- ggplot(data=heatmap_df, aes(x=ValueA, y=ValueB))+
    geom_tile(aes(fill=n))+
    theme_classic()+
    scale_x_continuous(breaks=c(1:6), limits=c(0,7))+
    scale_y_continuous(breaks=c(1:6), limits=c(0,7))+
    scale_fill_gradient(low = "#F0F0F0", high = "black", na.value="white", limits=c(1, maximum), breaks=c(1:maximum), labels=c(1:maximum))+
    ggtitle(paste("Condition", i ))
  #cat("  \n Condition:",  paste(i), "  \n")  
  print(figure2)
  #dev.off()
  cat("  \n")
}

```

We performed Chi-square tests to see if the reported values of our participants come from a uniform distribution. The results are shown in Table XXX. 
*Reported values in conditions with a dishonest partner (SD and CD) are significantly different from the uniform distribution, probably because participants cheated and matched player As' values, which in turn were sampled from a screwed distribution. Reported values in conditions with an honest partner (SH and CH) do not differ significantly from the uniform distributions, however, these results does not show whether participants cheated. The distribution can be uniform because they did not cheat or because they cheated by matching player As' values that in turn were sampled from a uniform distribution.*

ITT HIANYZIK A DF A TABLAZATBOL  

We've specified simulate.p.value = TRUE, so chisq.test will use a test statistic and P-value based on a Monte Carlo approach. In this case, there is no assumed chi-square distribution for the test statistic, so there is also no df parameter involved.  


```{r Chi-square}

probabilities <- c(rep(1/6, 6))

df.results <- data.frame(Condition=cond, df=NA, Chi.square=NA, p.value=NA)
for(i in cond)
{
  reported_B <- count(filter(rawdata, Condition==i), ValueB)$n
  results <- chisq.test(reported_B, p=probabilities, simulate.p.value = TRUE)
  
  df.results[which(df.results$Condition==i), ]$df <- results$parameter
  df.results[which(df.results$Condition==i), ]$Chi.square <- results$statistic
  df.results[which(df.results$Condition==i), ]$p.value <- results$p.value
  
  #cat("  \n Condition:",  paste(i), "  \n") 
  #print(results)

}

kable(df.results, digits=2)

```

## The number of doubles

The following figure shows the number of doubles reported by all participants as a function of time in each condition.
*It can be seen that the number of doubles does not increase over time.*

```{r Doubles and time}

```

The boxplots of Figure XXX show the distribution of the number of doubles in each condition, summed for all rounds of the game.

** FORD?TOTT SORRENDBEN VANNAK A BOXOK!*** a képaláírás egyezik, milyen sorrendben kéne lenni? tökre elfelejtettem ezeket, bocsi :(

```{r Box plots}

cond <- as.factor(unique(rawdata$Condition))
number.of.doubles <-  count(filter(rawdata, Double==TRUE), ID, Condition)

figure3 <- ggplot(data= number.of.doubles, aes(y=n, x=Condition))+
  geom_boxplot()+
  theme_minimal()+
  scale_x_discrete(name="Conditions")+
  scale_y_continuous(name='Number of doubles per 20 rounds')

#yourfilename=paste("figure3", ".jpg",sep="") #filename for sacing
#jpeg(file=yourfilename)
print(figure3)
#dev.off()

```

We tested, whether the number of doubles is higher than its expected value of 3.33 with Wilcoxon signed-rank U test, separately for each condition, see Table XXX.
*All tests showed that the number of doubles is significantly higher than expected by chance, meaning that participants cheated in all conditions.*

EBBE A T?BL?ZATBA MEG BELE KENE TENNI A MEDIAN NUMBER OF DOBLES-T, NEM? KÉSZ 

```{r Wilcoxon signed-rank U test}

w1 <- list()
median1 <- c()
for (i in groups$Condition) {
  current <- filter(participants, Condition == i)
  median1[i] <- median(current$Nbof_doubles)
  w1[[i]] <- wilcox.test(x=current$Nbof_doubles, mu=20/6, alternative = "greater")
}

results.wilcox.table1 <- data.frame(
  Condition = groups$Condition,
  p.value = c(w1[[1]]$p.value, w1[[2]]$p.value, w1[[3]]$p.value, w1[[4]]$p.value),
  W = c(w1[[1]]$statistic, w1[[2]]$statistic, w1[[3]]$statistic, w1[[4]]$statistic),
  median=c(median1))

kable(results.wilcox.table1, digits=10)

```

We compared the number of doubles in pairs of conditions with two-sample Wilcoxon tests (one-sided Mann-Whitney U tests). The results of testing the effect of dishonest partners vs honest partners in the simple game and in the charity game are shown in Table XXX.
*Both comparisons yielded significant results supporting the moral alignment hypothesis.*

```{r Mann-Whitney U test 1}

w2 <- list()
w2[[1]] <- wilcox.test(x=filter(participants, Condition=="SH")$Nbof_doubles, y=filter(participants, Condition=="SD")$Nbof_doubles, conf.int = TRUE, alternative = "less" )
w2[[2]] <- wilcox.test(x=filter(participants, Condition=="CH")$Nbof_doubles, y=filter(participants, Condition=="CD")$Nbof_doubles, conf.int = TRUE, alternative = "less" )

results.wilcox.table2 <- data.frame(
  Condition = c("SH-SD", "CH-CD"),
  p.value = c(w2[[1]]$p.value, w2[[2]]$p.value),
  W = c(w2[[1]]$statistic, w2[[2]]$statistic))

kable(results.wilcox.table2, digits=10)

```

The results of testing the effect of charity vs no charity with honest partner and with dishonest partner are shown in Table XXX.
*Both comparisons yielded significant results supporting the charity hypothesis*

```{r Mann-Whitney U test 2}

w3 <- list()
w3[[1]] <- wilcox.test(x=filter(participants, Condition=="SH")$Nbof_doubles, y=filter(participants, Condition=="CH")$Nbof_doubles, conf.int = TRUE, alternative = "less" )
w3[[2]] <- wilcox.test(x=filter(participants, Condition=="SD")$Nbof_doubles, y=filter(participants, Condition=="CD")$Nbof_doubles, conf.int = TRUE, alternative = "less" )

results.wilcox.table3 <- data.frame(
  Condition = c("SH-CH", "SD-CD"),
  p.value = c(w3[[1]]$p.value, w3[[2]]$p.value),
  W = c(w3[[1]]$statistic, w3[[2]]$statistic))

kable(results.wilcox.table3, digits=10)

```

Finally, we tested the effect of predictors with linear regression, see Table XXX. The dependent variable was the number of reported doubles (interval) and predictors were game (binary), partner (binary) and fingerratio (interval).

```{r}

#model1 <- lm(Nbof_doubles ~ Game + Partner + Fingerratio, data=participants)
#summary1 <- summary(model1)
#kable(summary1$coefficients, digits=3)

```

Residual standard error: `r #round(summary1$sigma, 3)`  
Multiple R-squared: `r #round(summary1$r.squared, 3)`  
Adjusted R-squared: `r #round(summary1$adj.r.squared, 3)`  
F-statistic: `r #round(summary1$fstatistic[[1]], 3)`  





